Logarithmic Space Verifiers on NP-complete

### 2019-10-31 doc/manuscript.tex
Improved the content doc
Fixed the proof

%TODO mandatory: add short abstract of the document
\begin{abstract}
P versus NP is considered as one of the most important open problems in computer science. This consists in knowing the answer of the following question: Is P equal to NP? A precise statement of the P versus NP problem was introduced independently by Stephen Cook and Leonid Levin. Since that date, all efforts to find a proof for this problem have failed. NP is the complexity class of languages defined by polynomial time verifiers M such that when the input is an element of the language with its certificate, then M outputs a string which belongs to a single language in P. Another major complexity classes are L and NL. The certificate-based definition of NL is based on logarithmic space Turing machine with an additional special read-once input tape: This is called a logarithmic space verifier. NL is the complexity class of languages defined by logarithmic space verifiers M such that when the input is an element of the language with its certificate, then M outputs 1. To attack the P versus NP problem, the NP-completeness is a useful concept. We demonstrate there is an NP-complete language defined by a logarithmic space verifier M such that when the input is an element of the language with its certificate, then M outputs a string which belongs to a single language in L.
\end{abstract}

@@

The $P$ versus $NP$ problem is a major unsolved problem in computer science \cite{CS00}. This is considered by many to be the most important open problem in the field \cite{CS00}. It is one of the seven Millennium Prize Problems selected by the Clay Mathematics Institute to carry a US\$1,000,000 prize for the first correct solution \cite{CS00}. It was essentially mentioned in 1955 from a letter written by John Nash to the United States National Security Agency \cite{AS17}. However, the precise statement of the $P=NP$ problem was introduced in 1971 by Stephen Cook in a seminal paper \cite{CS00}. In 2012, a poll of 151 researchers showed that 126 (83\%) believed the answer to be no, 12 (9\%) believed the answer is yes, 5 (3\%) believed the question may be independent of the currently accepted axioms and therefore impossible to prove or disprove, 8 (5\%) said either do not know or do not care or don't want the answer to be yes nor the problem to be resolved \cite{GA12}. It is fully expected that $P \neq NP$ \cite{Pap03}. Indeed, if $P = NP$ then there are stunning practical consequences \cite{Pap03}. For that reason, $P = NP$ is considered as a very unlikely event \cite{Pap03}. Certainly, $P$ versus $NP$ is one of the greatest open problems in science and a correct solution for this incognita will have a great impact not only in computer science, but for many other fields as well \cite{AS17}. Whether $P = NP$ or not is still a controversial and unsolved problem \cite{AS17}. In this work, we show some results that might be helpful in facing one of the most important open problems in computer science.

@@

@@

\begin{proof}
Given a valid instance $(\psi, K)$ for $MIN\oplus2SAT$ when $\psi$ has $m$ clauses, we can create a certificate array $A$ which contains $K$ different natural numbers in ascending order which represents the indexes of the clauses in $\psi$ that we are going to remove from the instance. We read at once the elements of the array $A$ and we reject whether this is not a valid certificate: That is when the numbers are not sorted in ascending order, or the array $A$ does not contain exactly $K$ elements, or the array $A$ contains a number that is not between $1$ and $m$. While we read the elements of the array $A$, we remove the clauses from the instance $(\psi, K)$ for $MIN\oplus2SAT$ just creating another instance $\phi$ for $\textit{XOR 2SAT}$ where the Boolean formula $\phi$ does not contain the $K$ different indexed clauses $\psi$ represented by the numbers in $A$. Therefore, we obtain that:
\[(\psi, K) \in MIN\oplus2SAT \textit{ if and only if } \phi \in \textit{XOR 2SAT}.\]
Furthermore, we can make this verification in logarithmic space such that the array $A$ is placed on the special read-once tape, because we read at once the elements in the array $A$ and we assume the clauses in the input $\psi$ are indexed from left to right. Hence, we only need to iterate from the elements of the array $A$ to verify whether the array is a valid certificate and also remove the $K$ different clauses from the Boolean formula $\psi$ when we write the final clauses to the output. This logarithmic space verification will be the Algorithm \ref{2-algorithm}. We assume whether a value does not exist in the array $A$ into the cell of some position $i$ when $A[i] = undefined$. In addition, we reject immediately when the following comparisons
\[A[i] \leq max \vee A[i] < 1 \vee A[i] > m\]
hold at least into one single binary digit. Note, in the loop $j$ from $min$ to $max - 1$, we do not output any clause when $max - 1 < min$.

@@

@@

This work is implemented into a GitHub Project programmed in Scala \cite{FV19}. In this GitHub Project, we use the Assertion on the properties of the instances of each problem and the Unit Test for checking the correctness of every reduction \cite{FV19}. We need to install JDK 8 in order to test the Scala Project \cite{OD16}. In addition, we need to install SBT to run the unit test (we could run the unit test with the $\textbf{sbt test}$ command) \cite{OD16}.

>>>>

%TODO mandatory: add short abstract of the document
\begin{abstract}
P versus NP is considered as one of the most important open problems in computer science. This consists in knowing the answer of the following question: Is P equal to NP? A precise statement of the P versus NP problem was introduced independently by Stephen Cook and Leonid Levin. Since that date, all efforts to find a proof for this problem have failed. NP is the complexity class of languages defined by polynomial time verifiers M such that when the input is an element of the language with its certificate, then M outputs a string which belongs to a single language in P. Another major complexity classes are L and NL. The certificate-based definition of NL is based on logarithmic space Turing machine with an additional special read-once input tape: This is called a logarithmic space verifier. NL is the complexity class of languages defined by logarithmic space verifiers M such that when the input is an element of the language with its certificate, then M outputs 1. To attack the P versus NP problem, the NP-completeness is a useful concept. We demonstrate there is an NP-complete language defined by a logarithmic space verifier M such that when the input is an element of the language with its certificate, then M outputs a string which belongs to a single language in L. In this way, we obtain that L = NL and P = NP cannot be both true or false at the same time and thus, the complexity class L is not equal to NP.
\end{abstract}

@@

In previous years there has been great interest in the verification or checking of computations \cite{LIP90}. Interactive proofs introduced by Goldwasser, Micali and Rackoff and Babi can be viewed as a model of the verification process \cite{LIP90}. Dwork and Stockmeyer and Condon have studied interactive proofs where the verifier is a space bounded computation instead of the original model where the verifer is a time bounded computation \cite{LIP90}. In addition, Blum and Kannan has studied another model where the goal is to check a computation based solely on the final answer \cite{LIP90}. More about probabilistic logarithmic space verifiers have been shown on a technique of Lipton \cite{LIP90}. In this work, we show some results about the logarithmic space verifiers applied to the class $NP$ which solve one of the most important open problems in computer science, that is $L$ versus $NP$.

\section{Motivation}

The $P$ versus $NP$ problem is a major unsolved problem in computer science \cite{CS00}. This is considered by many to be the most important open problem in the field \cite{CS00}. It is one of the seven Millennium Prize Problems selected by the Clay Mathematics Institute to carry a US\$1,000,000 prize for the first correct solution \cite{CS00}. It was essentially mentioned in 1955 from a letter written by John Nash to the United States National Security Agency \cite{AS17}. However, the precise statement of the $P=NP$ problem was introduced in 1971 by Stephen Cook in a seminal paper \cite{CS00}. In 2012, a poll of 151 researchers showed that 126 (83\%) believed the answer to be no, 12 (9\%) believed the answer is yes, 5 (3\%) believed the question may be independent of the currently accepted axioms and therefore impossible to prove or disprove, 8 (5\%) said either do not know or do not care or don't want the answer to be yes nor the problem to be resolved \cite{GA12}. It is fully expected that $P \neq NP$ \cite{Pap03}. Indeed, if $P = NP$ then there are stunning practical consequences \cite{Pap03}. For that reason, $P = NP$ is considered as a very unlikely event \cite{Pap03}. Certainly, $P$ versus $NP$ is one of the greatest open problems in science and a correct solution for this incognita will have a great impact not only in computer science, but for many other fields as well \cite{AS17}. Whether $P = NP$ or not is still a controversial and unsolved problem \cite{AS17}. We show some results that are a breakthrough in the future path of solving this outstanding problem.

@@

From the early days of automata and complexity  theory,  two  different  models  of  Turing  machines  are  considered, the offline and online machines \cite{KU15}. Each model has a read-only input tape and some work tapes \cite{KU15}. The offline machines may read their input two-way while the online machines are not allowed to move the input head to the left \cite{KU15}. In the terminology of the (generalized) Turing machine models are called two-way and one-way Turing machines respectively \cite{KU15}.

Hartmanis and Mahaney have investigated the classes $1L$ and $1NL$ of languages recognizable by deterministic one-way logarithmic space Turing machine and nondeterministic one-way logarithmic space Turing machine, respectively \cite{HM81}. They have shown that $1L \neq 1NL$ (by looking at a uniform variant of the string non-equality problem from communication complexity theory) and have defined a natural complete problem for $1NL$ under deterministic one-way logarithmic space reductions \cite{HM81}. Furthermore, they have proven that $1NL \subseteq L$ if and only if $L=NL$ \cite{HM81}.

\begin{theorem}
\label{argument}
If the Hypothesis \ref{hypothesis} is true, then $L = NL$ and $P = NP$ cannot be both true or false at the same time and thus, $L \neq NP$.
\end{theorem}

\begin{proof}
We can simulate the computation $M(w, u) = y$ in the Hypothesis \ref{hypothesis} by a nondeterministic logarithmic space Turing machine $N$, such that $N(w) = y$ since we can read the certificate string $u$ within the read-once tape by a work tape in a nondeterministic logarithmic space generation of symbols contained in $u$ \cite{Pap03}. Certainly, we can simulate the reading of one symbol from the string $u$ into the read-once tape just nondeterministically generating the same symbol in the work tapes using a logarithmic space \cite{Pap03}.

If we suppose that $L \subset 1NL$, then we can accept the elements of the language $L_{1} \in L$ by a nondeterministic one-way logarithmic space Turing machine $M'$. In this way, there is a nondeterministic logarithmic space Turing machine $M''(w) = M'(N(w))$ which will output 1 when $w \in L_{2}$. Consequently, $M''$ is a nondeterministic logarithmic space Turing machine which decides the language $L_{2}$. The reason is because we can simulate the output string of $N(w)$ within a read-once tape and thus, we can compute in a nondeterministic logarithmic space the logarithmic space composition using the same techniques of the logarithmic space composition reduction, but without any reset of the computation \cite{Pap03}. Certainly, we do not need to reset the computation of $N(w)$ for the reading at once of a symbol in the output string of $N(w)$ by the nondeterministic one-way logarithmic space Turing machine $M'$. Therefore, $L_{2}$ is in $NL$ and thus, $L_{2} \in P$ due to $NL \subseteq P$ \cite{Pap03}. If any single $\textit{NP--complete}$ problem can be solved in polynomial time, then $P = NP$ \cite{CLRS01}. Since $L_{2} \in P$ and $L_{2} \in \textit{NP--complete}$, then we obtain the complexity class $P$ is equal to $NP$ under the assumption of $L \subset 1NL$.

Hartmanis and Mahaney have also shown with their result that if $1NL \subseteq L$ or even $1NL \subset L$, then $L=NL$, because they proved there is a complete problem for both $1NL$ and $NL$ at the same time \cite{HM81}. If this way, if $L \neq NL$, then $L \subset 1NL$ by contraposition \cite{Pap03}. Since we already obtained that $P = NP$ under the assumption of $L \subset 1NL$, therefore if $L \neq NL$, then $P = NP$. Hence, if $P \neq NP$, then $L = NL$ by contraposition \cite{Pap03}. Consequently, $L = NL$ and $P = NP$ cannot be both true or false at the same time and thus, $L \neq NP$.
\end{proof}

@@

\begin{proof}
Given a valid instance $(\psi, K)$ for $MIN\oplus2SAT$ when $\psi$ has $m$ clauses, we can create a certificate array $A$ which contains $K$ different natural numbers in ascending order which represents the indexes of the clauses in $\psi$ that we are going to remove from the instance. We read at once the elements of the array $A$ and we reject whether this is not a valid certificate: That is when the numbers are not sorted in ascending order, or the array $A$ does not contain exactly $K$ elements, or the array $A$ contains a number that is not between $1$ and $m$. While we read the elements of the array $A$, we remove the clauses from the instance $(\psi, K)$ for $MIN\oplus2SAT$ just creating another instance $\phi$ for $\textit{XOR 2SAT}$ where the Boolean formula $\phi$ does not contain the $K$ different indexed clauses $\psi$ represented by the numbers in $A$. Therefore, we obtain the array $A$ should be valid according to the Theorem \ref{proof} when:
\[(\psi, K) \in MIN\oplus2SAT \textit{ if and only if } \phi \in \textit{XOR 2SAT}.\]
Furthermore, we can make this verification in logarithmic space such that the array $A$ is placed on the special read-once tape, because we read at once the elements in the array $A$ and we assume the clauses in the input $\psi$ are indexed from left to right. Hence, we only need to iterate from the elements of the array $A$ to verify whether the array is a valid certificate and also remove the $K$ different clauses from the Boolean formula $\psi$ when we write the final clauses to the output. This logarithmic space verification will be the Algorithm \ref{2-algorithm}. We assume whether a value does not exist in the array $A$ into the cell of some position $i$ when $A[i] = undefined$. In addition, we reject immediately when the following comparisons
\[A[i] \leq max \vee A[i] < 1 \vee A[i] > m\]
hold at least into one single binary digit. Note, in the loop $j$ from $min$ to $max - 1$, we do not output any clause when $max - 1 < min$.

@@

\begin{theorem}
$L = NL$ and $P = NP$ cannot be both true or false at the same time and thus, $L \neq NP$.
\end{theorem}

\begin{proof}
This is a direct consequence of Theorems \ref{argument} and \ref{result}.
\end{proof}

@@

This work is implemented into a Project programmed in Scala \cite{FV19}. In this Project, we use the Assertion on the properties of the instances of each problem and the Unit Test for checking the correctness of every reduction \cite{FV19}. We need to install JDK 8 in order to test the Scala Project \cite{OD16}. In addition, we need to install SBT to run the unit test (we could run the unit test with the $\textbf{sbt test}$ command) \cite{OD16}.
